G\"{o}del's Ontological Proof of God's Existence (Draft)

### 2013-08-25 Natural Deduction/GodProof-ND.tex
Add remarks about box elimination
Improve ND proof

@@

\begin{theorem}
\label{T1}
Positive properties are possibly exemplified:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$
\end{theorem}
\begin{proof} \hfill
\begin{prooftree}
        \AXC{Axiom \ref{A2}} \dashedLine
        \UIC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
        \UIC{$ \all \psi.[(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
        \UIC{$(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \neg \varphi'(x)]) \imp P(\neg \varphi')$} \doubleLine
        \UIC{$(P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp P(\neg \varphi')$}
                        \AXC{Axiom \ref{A1}} \dashedLine
                        \UIC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$} \RightLabel{$\all_E$}
                        \UIC{$ P(\neg \varphi') \biimp \neg P(\varphi') $} \doubleLine
                 \BIC{$ (P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp \neg P(\varphi') $} \doubleLine
                 \UIC{$ P(\varphi') \imp \pos \ex x.\varphi'(x) $} \RightLabel{$\all_I$}
                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
\end{prooftree}
\end{proof}

@@

ToDo: instead of using the $\nec_E$ rule, we should use the $M$ axiom.

@@

\noindent
Let the following derivation with the open assumption $G(x)$ be $\Pi_2[G(x)]$: 
\begin{prooftree}
       \AXC{$\psi(x)^3$}
             	\AXC{$\Pi_1[G(x)] $}  \dashedLine
		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp_E$}
       		\BIC{$P(\psi)$} 
		              \AXC{Axiom \ref{A4}} \dashedLine
       	          \UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\all_E$}  
		              \UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp_E$}  
           \BIC{$\Box P(\psi)$}\RightLabel{$\imp_I^3$} 
           \UIC{$\psi(x)\imp\Box P(\psi)$} 
\end{prooftree}

@@

\noindent
Let the following derivation with the open assumption $G(x)$ be $\Pi_4[G(x)]$: 
\begin{prooftree}
       \AXC{$\psi(x)^6 $} \RightLabel{}
             	\AXC{$\Pi_2$}  \dashedLine
		\UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp_E$}
		\BIC{$\Box P(\psi)$}
		  	\AXC{$\Box P(\psi)^7 $}\RightLabel{$\Box_E$}  
       			\UIC{$P(\psi)$}
				\AXC{$\Pi_3$}  \dashedLine\RightLabel{}
       				\UIC{$P(\psi)\imp \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_E$}  
       			\BIC{$\all x .(G(x)\imp \psi(x))$}\RightLabel{Necessitation}
			\UIC{$\Box \all x .(G(x)\imp \psi(x))$}  \RightLabel{$\imp_I^7$} 
			\UIC{$\Box P(\psi)\imp \Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_E$} 
         	 \BIC{$\Box \all x .(G(x)\imp \psi(x)$}\RightLabel{$\imp_I^6$}
           \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}  
\end{prooftree}

@@

\begin{prooftree}
       \AXC{$G(x)^8 $} \RightLabel{}
             	\AXC{$\Pi_4[G(x)^8]$}  \dashedLine
		\UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\all_I$}
		\UIC{$\all \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\RightLabel{$\wedge_I$}
		\BIC{$G(x)\wedge \forall \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{D\ref{D2}} 
       			\UIC{$\ess{G}{x}$} \RightLabel{$\imp_I^8$}
			\UIC{$G(x)\imp \ess{G}{x}$}
			  \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
\end{prooftree}
\end{proof}

@@

\begin{definition}
\label{D3}
\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
$$
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)]
$$
\end{definition}

>>>>

ToDo: Investigate the relationship between the $\nec_E$ rule and the $M$ axiom (reflexivity). We don't want $M$ to be provable in our calculus.

@@

\begin{theorem}
\label{T1}
Positive properties are possibly exemplified:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$
\end{theorem}
\begin{proof} \hfill
\begin{prooftree}
        \AXC{Axiom \ref{A2}} \dashedLine
        \UIC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
        \UIC{$ \all \psi.[(P(\rho) \wedge \nec \all x.[\rho(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
        \UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \doubleLine
        \UIC{$(P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp P(\neg \rho)$}
                        \AXC{Axiom \ref{A1}} \dashedLine
                        \UIC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$} \RightLabel{$\all_E$}
                        \UIC{$ P(\neg \rho) \biimp \neg P(\rho) $} \doubleLine
                 \BIC{$ (P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp \neg P(\rho) $} \doubleLine
                 \UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $} \RightLabel{$\all_I$}
                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
\end{prooftree}
\end{proof}

@@

@@

\noindent
Let the following derivation with the open assumption $G(x)$ be $\Pi_2[G(x)]$: 
\begin{prooftree}
       \AXC{$\psi(x)^3$}
             	\AXC{$\Pi_1[G(x)] $}  \dashedLine
		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp_E$}
       		\BIC{$P(\psi)$} 
		              \AXC{Axiom \ref{A4}} \dashedLine
       	          \UIC{$\all \varphi .(P(\varphi)\imp \Box P(\varphi))$}\RightLabel{$\all_E$}  
		              \UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp_E$}  
           \BIC{$\Box P(\psi)$}\RightLabel{$\imp_I^3$} 
           \UIC{$\psi(x)\imp\Box P(\psi)$} 
\end{prooftree}

@@

\noindent
Let the following derivation with the open assumption $G(x)$ be $\Pi_4[G(x)]$: 
\begin{prooftree}
       \AXC{$\psi(x)^6 $} \RightLabel{}
             	\AXC{$\Pi_2$}  \dashedLine
		\UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp_E$}
		\BIC{$\Box P(\psi)$}
		  	\AXC{$\Box P(\psi)^7 $}\RightLabel{$\Box_E$}  
       			\UIC{$P(\psi)$}
				\AXC{$\Pi_3$}  \dashedLine\RightLabel{}
       				\UIC{$P(\psi)\imp \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_E$}  
       			\BIC{$\all x .(G(x)\imp \psi(x))$}\RightLabel{Necessitation}
			\UIC{$\Box \all x .(G(x)\imp \psi(x))$}  \RightLabel{$\imp_I^7$} 
			\UIC{$\Box P(\psi)\imp \Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_E$} 
         	 \BIC{$\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_I^6$}
           \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}  
\end{prooftree}

@@

\begin{prooftree}
       \AXC{$G(x)^8 $} \RightLabel{}
             	\AXC{$\Pi_4[G(x)^8]$}  \dashedLine
		\UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\all_I$}
		\UIC{$\all \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\RightLabel{$\wedge_I$}
		\BIC{$G(x)\wedge \forall \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{D\ref{D2}} 
       			\UIC{$\ess{G}{x}$} \RightLabel{$\imp_I^8$}
			\UIC{$G(x)\imp \ess{G}{x}$} \RightLabel{$\all_I$}
			  \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
\end{prooftree}
\end{proof}

@@

\begin{definition}
\label{D3}
\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
$$
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
$$
\end{definition}
